---
title: "Untitled"
output:
  html_document: default
  pdf_document: default
date: "2024-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#question 2a
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
leaddata<-read.csv("/Users/helenwolde/Desktop/consulting/consulting/DataRaw/lead-iq-01.csv")

leaddata[10,2]<-99 #changing 999 from 99

ggplot(leaddata, aes(x = Smelter, y = IQ, fill = Smelter)) +
  geom_boxplot() +
  ylim(40, 1000) +  # Adjust y-axis limits
  labs(title = "IQ Levels by Location", x = "Location", y = "IQ Level") +
  theme_minimal()


```

#question2b
```{r}
library(knitr)
library(dplyr)
library(pander)
df_summary <- leaddata %>%
  group_by(Smelter) %>%
  summarize(
    Mean_IQ = mean(IQ, na.rm = TRUE),
    SD_IQ = sd(IQ, na.rm =TRUE)
  )
pander(df_summary, caption = "Mean and Standard deviation of IQ by Location")
```

The mean IQ score in the Far smelter location is 17.9 points higher than that of the Near location. This difference may suggest that environmental, socioeconomic, or other factors associated with the locations could be impacting IQ levels.

The standard deviation for the Far location is 111.9, which appears unusually high and may suggest that there is significant variability in IQ scores among individuals in this group.

The standard deviation for the Near location is 12.17, indicating that the IQ scores are more closely clustered around the mean compared to the Far location.

#Question 2c
```{r}
summary_stats <- leaddata %>%
  group_by(Smelter) %>%
  summarize(
    Mean_IQ = mean(IQ),
    SD_IQ = sd(IQ)
  )
print(summary_stats)
```

#2d In-line calculations
```{r}
far_iq<-leaddata %>% filter(Smelter=="Far") %>% select(IQ)
sum(far_iq)/nrow(far_iq)

near_iq<-leaddata %>% filter(Smelter=="Near") %>% select(IQ)
sum(near_iq)/nrow(near_iq)

```



